Zero-shot task adaptation by homoiconic meta-mapping

### 2019-09-07 writing/embedded_meta_learning_main.tex
Tweaking figure, trying to cut length

\begin{document}
\title{Zero-shot adaptation by meta-mapping between tasks}
\author{%
Andrew K. Lampinen\\
Department of Psychology\\
Stanford University\\
\texttt{lampinen@stanford.edu}\\
\And
James L. McClelland\\
Department of Psychology\\
Stanford University\\
\texttt{mcclelland@stanford.edu}\\
}
\date{}
\maketitle

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks zero-shot. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from functional programming and recent work in meta-learning to propose a class of Embedded Meta-Learning (EML) architectures that represent both data and tasks in a shared latent space. EML architectures are applicable to any type of machine learning task, including supervised learning and reinforcement learning. We demonstrate the utility of this perspective by exhibiting zero-shot remapping of behavior to adapt to new tasks.
\end{abstract}

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose using architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recursively treat its task representations as data points that can be transformed to produce new task representations, our approach is able to achieve this flexibility parsimoniously, without adding additional parameters beyond what we use for basic meta-learning. We suggest that approaches like ours will be key to building more intelligent and flexible deep learning systems. \par

\section{Meta-mapping}
\vspace{-0.5em} % section
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. (We discuss the relationship between our work and prior work in section \ref{sec_discussion}.) \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to some variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way to achieve this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose, or trying to achieve some orthogonal goal. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), our architecture can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
There are different ways this adaptation can be cued, for example by examples of similar task mappings (``you've played winning and losing poker, now try to lose at blackjack'') or simply by natural-language instructions (``try to lose at blackjack'', see appendix \ref{app_why_meta_mapping} for a comparison to conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\section{Embedded Meta-Learning (EML) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system that embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for individual data points and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system can learn to perform meta-mappings to follow instructions or change behavior. That is, it can transform task representations using the same components it uses to transform basic data points, because the task representations and basic data points are embedded in the same space. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, as we will show, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-task distribution. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
%% basic meta learning
\begin{scope}[shift={(0.4, 0.4)}]
\draw[boundingbox, fill=white] (-3, -4.3) rectangle (2.5, 3.2);
\end{scope}
\begin{scope}[shift={(0.2, 0.2)}]
\draw[boundingbox, fill=white] (-3, -4.3) rectangle (2.5, 3.2);
\end{scope}

@@

\node [block] at (-1.5, -1.75) (M) {\(\mathcal{M}\)};
\path [arrow] ([yshift=-5]D1.north) to (M);

\node at (-1, -0.7) (zfunc) {\(z^{func}\)};
\path [arrow, out=90, in=-135] (M.north) to ([xshift=6,yshift=3]zfunc.south west);

\node[block] at (0, -0.25) (H) {\(\mathcal{H}\)};
\path [arrow, out=45, in=180] ([yshift=-3]zfunc.north) to (H.west);

\node [block, dashed] at (1.5, -0.25) (F) {\(F_{z^{func}}\)};
\path[arrow] (H.east) to (F.west);

@@

\node at (1.5, 1.75) (outputs) {
%\(z^{pred}_{out}\)};
\(\left\{
\begin{matrix}
z^{pred}_{out,0},\\
$\vdots$
\end{matrix}\right\}\)};
\node at (1.5, 2.65) (predictions) {Predictions};

@@

\node [align=center, text width=1.25 cm] at (-1.5, 1.75) (dispatch) {\baselineskip=12pt Loss or eval or ...\par};

\path [arrow, dashed] (outputs.west) to ([xshift=-3]dispatch.east);

%% meta mapping
\begin{scope}[shift={(6.4, 2)}]
\draw[boundingbox, draw=bpurp, fill=white] (-3, -4.3) rectangle (2.6, 3.2);
\node[bpurp] at (-1.85, 2.9) {Meta-mapping};
\node at (-1.5, -4) (metaexamples) {Examples};
\node at (-1.5, -3.2) (metaD1) {
\(\left\{
\begin{matrix}
(z^{func}_{in,0}, z^{func}_{targ,0})\\
$\vdots$
\end{matrix}\right\}\)};

@@

\node [block] at (-1.5, -1.75) (metaM) {\(\mathcal{M}\)};
\path [arrow] ([yshift=-5]metaD1.north) to (metaM);

\node at (-1, -0.7) (metazfunc) {\(z^{meta}\)};
\path [arrow, out=90, in=-135] (metaM.north) to ([xshift=6,yshift=3]metazfunc.south west);

\node[block] at (0, -0.25) (metaH) {\(\mathcal{H}\)};
\path [arrow, out=45, in=180] ([yshift=-3]metazfunc.north) to (metaH.west);

\node [block, dashed] at (1.5, -0.25) (metaF) {\(F_{z^{meta}}\)};
\path[arrow] (metaH.east) to (metaF.west);

@@

\node at (1.5, 1.75) (metaoutputs) {
%\(z^{pred}_{out}\)};
\(\left\{
\begin{matrix}
z^{func,pred}_{out,0},\\
$\vdots$
\end{matrix}\right\}\)};
\node at (1.5, 2.65) (metapredictions) {Predictions};

@@

\node [align=center, text width=1.5 cm] at (-1.5, 1.75) (metadispatch) {Loss (training)};
\path [arrow, dashed] (metaoutputs.west) to ([xshift=-3]metadispatch.east);
\end{scope}
\path [arrow, draw=bpurp, dotted, out=-40, in=-140] ([xshift=-1, yshift=3]zfunc.south) to ([xshift=19, yshift=3]metaD1.west);

%% evaluating meta mapping
\begin{scope}[shift={(11.8, 0)}]
\draw[boundingbox, draw=bblue, fill=white] (-2.25, -4.3) rectangle (2.5, 3.2);
\node[bblue, text width=2.5cm] at (-0.9, 2.75) {Meta-mapping evaluation};

@@

\node at (-1, -0.7) (metaevalzfunc) {\(z^{func,pred}\)};

\node[block] at (0, -0.25) (metaevalH) {\(\mathcal{H}\)};
\path [arrow, out=45, in=180] ([yshift=-3]metaevalzfunc.north) to (metaevalH.west);

\node [block, dashed] at (1.5, -0.25) (metaevalF) {\(F_{z^{func,pred}}\)};
\path[arrow] (metaevalH.east) to (metaevalF.west);

@@

\node at (1.5, 1.75) (metaevaloutputs) {
%\(z^{pred}_{out}\)};
\(\left\{
\begin{matrix}
z^{pred}_{out,0},\\
$\vdots$
\end{matrix}\right\}\)};
\node at (1.5, 2.65) (metaevalpredictions) {Predictions};

@@

\node [align=center, text width=1.25 cm] at (-1.5, 1.75) (metaevaldispatch) {Loss (eval)};
\path [arrow, dashed] (metaevaloutputs.west) to ([xshift=-3]metaevaldispatch.east);

@@

\caption{The EML architecture allows for transformations at different levels of abstraction. Basic meta-learning proceeds as in the gray box: a dataset consisting of (input, output) tuples is processed by the meta-network \(\mathcal{M}\) to produce a function embedding \(z^{func}\), which is processed by the hyper network \(\mathcal{H}\) to parameterize a function \(F_{z^{func}}\), which attempts to compute the transformation on held-out probe inputs. However, our approach goes byeond this basic meta-learning step. The function embedding \(z^{func}\) can then be seen as a single input or output at the next level of abstraction, when the same networks \(\mathcal{M}\) and \(\mathcal{H}\) are used to transform function embeddings based on examples of a meta-mapping. To evaluate meta-mapping performance, a probe embedding from a held-out funciton is transformed by the architecture to yield a predicted embedding for the transformed task. The performance of this predicted embedding is evaluated by moving back down a level of abstraction and evaluating on the actual target task. Because the function embedding is predicted by a transformation rather than computed from examples, this allows for zero-shot performance of a new task. (Blocks with solid edges denote learnable deep networks, and $F_{z}$ is a deep network parameterized by $\mathcal{H}$ conditioned on function embedding \(z\). See the text and appendix \ref{app_model_details} for further details.)} \label{architecture_inference_fig}

@@

More formally, we take a \emph{functional} perspective on learning. From this perspective, the data points that one function receives can themselves be functions\footnote{Indeed, any data point can be represented as a constant function that outputs the data point.}. The key insight is that then our architecture can transform data points\footnote{Where ``data'' is a quite flexible term. The approach is agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.} to perform basic tasks (as is standard in machine learning), but it can also transform these task functions to adapt to new tasks. This is related to the idea of a homoiconicity, defined above. Under this perspective, basic tasks and meta-mappings from task to task are really the same type of problem. The functions at one level of abstraction (the basic tasks) become inputs and outputs for higher-level functions at the next level of abstraction (meta-mapping between tasks). \par
Specifically, we embed each input, target, or mapping into a shared representational space $Z$. This means that single data points are embedded in the same space as the representation of a functionr or an entire dataset. See appendix \ref{app_lesion_results_shared_z} for discussion of why $Z$ should be shared. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Model outputs are decoded from $Z$ by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Target outputs are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$.\par

@@

More explicitly, suppose we have a dataset of example input, target pairs ($D_1 = \{(x_0, y_0), ...\}$), and some input $x$ from a probe dataset $D_2$ for which to predict an output $\hat{y}$. $\hat{y}$ would be generated as follows: 

@@

where $F_{z^{func}}$ is the meta-learners guess at a representation for the function underlying the training dataset $D_1$:

@@

Alternatively, we could map from language to a meta-mapping embedding, rather than inducing it from examples of the meta-mapping. This corresponds to the human ability to change behavior in response to instructions. The key feature of our architecture -- the fact that tasks, data, and language are all embedded in a shared space -- allows for substantial flexibility within a unified framework. Furthermore, our approach is parsimonious. Because it uses the same meta-learner for both basic tasks and meta-mappings, this increased flexibility does not require any added parameters.\footnote{At least in principle, in practice of course increasing network size might be more beneficial for EML architectures performing meta-mappings as well as basic tasks than for those performing only basic tasks.}

%\subsection{Boolean functions}
%As a proof of concept, we first evaluated the system on a simple task of computing boolean functions on boolean inputs. Specifically, we can imagine mapping from binary-valued input vectors to a single binary output representing the evaluation of some logical proposition on the input. We train the system to figure out how to compute the predicate from seeing a subset of the (input, output) pairs. \par   
%We can then train the system to do various meta-mappings, for example identifying certain types of functions (like XOR) or negating predicates. \par
%
\section{Learning multivariate polynomials} \label{sec_poly}
\vspace{-0.5em} % section
\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/basic_results.png}
\caption{The polynomials domain, section \ref{sec_poly}.}
\label{poly_basic_results}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning.png}
\caption{The cards domain, section \ref{sec_cards}.}
\label{cards_basic_results}
\end{subfigure}%
\caption{The EML system succeeds at the basic meta-learning tasks, which are a necessary prerequisite for meta-mappings. (\subref{poly_basic_results}) Basic meta-learning in the polynomials domain, section \ref{sec_poly}. The system successfully generalizes to held out polynomials. The solid line indicates optimal performance; the dashed line indicates chance performance (an untrained model). (\subref{cards_basic_results}) Basic meta-learning in the cards domain, section \ref{sec_cards}. The system successfully generalizes to held out games, both when trained on a random sample of half the tasks, or when a targeted subset is held out. The gray dashed line indicates chance performance, while the solid lines are optimal performance. The orange dashed lines show how well the system could perform by playing the strategy from the most correlated trained task. The fact that it generally exceeds this difficult baseline shows a deeper form of generalization than just memorizing a few strategies and picking the closest. Error-bars throughout are bootstrap 95\%-confidence intervals, numerical values for plots can be found in appendix \ref{app_numerical_results}.}
\end{figure}

>>>>

\begin{document}
\title{Zero-shot task adaptation by homoiconic meta-mapping}
\author{%
Andrew K. Lampinen\\
Department of Psychology\\
Stanford University\\
\texttt{lampinen@stanford.edu}\\
\And
James L. McClelland\\
Department of Psychology\\
Stanford University\\
\texttt{mcclelland@stanford.edu}\\
}
\date{}
\maketitle

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks zero-shot. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from functional programming and recent work in meta-learning to propose a class of Homoiconic Meta-Mapping (HoMM) approaches that represent data points and tasks in a shared latent space, and learn to infer transformations of that space. HoMM approaches can be applied to any type of machine learning task, including supervised learning and reinforcement learning. We demonstrate the utility of this perspective by exhibiting zero-shot remapping of behavior to adapt to new tasks.
\end{abstract}

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose using architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recursively treat its task representations as data points, and transform them to produce new task representations, our approach is able to achieve this flexibility parsimoniously. We suggest that approaches like ours will be key to building more intelligent and flexible deep learning systems. \par

\vspace{-0.25em}
\section{Meta-mapping}
\vspace{-0.5em} % section
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. (We discuss the relationship between our work and prior work in section \ref{sec_discussion}.) \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to some variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way to achieve this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose, or trying to achieve some orthogonal goal. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), our architecture can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
This adaptation can be cued in different ways, for example by examples of similar task mappings (after winning and losing at poker, try to lose at blackjack) or simply by natural-language instructions (``try to lose at blackjack'', see appendix \ref{app_why_meta_mapping} for a comparison to conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step toward more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\vspace{-0.25em}
\section{Homoiconic meta-mapping (HoMM) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose HoMM architectures, composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item Input/output systems: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system that embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior, and to systematically transform task embeddings.
\end{enumerate}
\vspace{-0.75em}
These architectures are homoiconic because they have a completely shared space $Z$ for individual data points and tasks. Why is this useful? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system can learn to perform meta-mappings to follow instructions or change behavior. That is, it can transform task representations using the same components it uses to transform basic data points. (See appendix \ref{app_lesion_results_shared_z} for further discussion.) \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, as we will show, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-mapping distribution. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
%% basic meta learning
\begin{scope}[shift={(0.4, 0.4)}]
\draw[boundingbox, fill=white] (-3, -4.3) rectangle (2.5, 3.2);
\end{scope}
\begin{scope}[shift={(0.2, 0.2)}]
\draw[boundingbox, fill=white] (-3, -4.3) rectangle (2.5, 3.2);
\end{scope}

@@

\node [block] at (-1.5, -1.95) (M) {\(\mathcal{M}\)};
\path [arrow] ([yshift=-5]D1.north) to (M);

\node at (-1, -1) (zfunc) {\(z^{func}\)};
\path [arrow, out=90, in=-135] (M.north) to ([xshift=6,yshift=3]zfunc.south west);

\node[block] at (0, -0.55) (H) {\(\mathcal{H}\)};
\path [arrow, out=45, in=180] ([yshift=-3]zfunc.north) to (H.west);

\node [block, dashed] at (1.5, -0.55) (F) {\(F_{z^{func}}\)};
\path[arrow] (H.east) to (F.west);

@@

\node at (1.5, 1) (outputs) {
%\(z^{pred}_{out}\)};
\(\left\{
\begin{matrix}
z^{pred}_{out,0},\\
$\vdots$
\end{matrix}\right\}\)};
\node at (1.5, 1.75) (predictions) {Predictions};

@@

\node at (-1.5, 1.75) (probetargs) {Probe targets};
\node at (-1.5, 1) (D2targs) {
%\(z^{prb}_{in}\)};
\(\left\{
\begin{matrix}
z^{prb}_{targ,0}\\
$\vdots$
\end{matrix}\right\}\)};

\node [align=center, text width=1.25 cm] at (0, 2.25) (dispatch) {\baselineskip=12pt Loss\par};

\path [arrow, dashed, out=180, in=-90] ([xshift=3]outputs.west) to (dispatch.south);

\path [arrow, dashed, out=0, in=-90] ([xshift=-3.5]D2targs.east) to (dispatch.south);

%% meta mapping
\begin{scope}[shift={(6.4, 1.5)}]
\draw[boundingbox, draw=bpurp, fill=white] (-3, -4.3) rectangle (2.6, 3.2);
\node[bpurp] at (-1.85, 2.9) {Meta-mapping};
\node at (-1.5, -4) (metaexamples) {Examples};
\node at (-1.5, -3.2) (metaD1) {
\(\left\{
\begin{matrix}
(z^{func}_{in,0}, z^{func}_{targ,0})\\
$\vdots$
\end{matrix}\right\}\)};

@@

\node [block] at (-1.5, -1.95) (metaM) {\(\mathcal{M}\)};
\path [arrow] ([yshift=-5]metaD1.north) to (metaM);

\node at (-1, -1) (metazfunc) {\(z^{meta}\)};
\path [arrow, out=90, in=-135] (metaM.north) to ([xshift=6,yshift=3]metazfunc.south west);

\node[block] at (0, -0.55) (metaH) {\(\mathcal{H}\)};
\path [arrow, out=45, in=180] ([yshift=-3]metazfunc.north) to (metaH.west);

\node [block, dashed] at (1.5, -0.55) (metaF) {\(F_{z^{meta}}\)};
\path[arrow] (metaH.east) to (metaF.west);

@@

\node at (1.5, 1) (metaoutputs) {
%\(z^{pred}_{out}\)};
\(\left\{
\begin{matrix}
z^{func,pred}_{out,0},\\
$\vdots$
\end{matrix}\right\}\)};
\node at (1.5, 1.75) (metapredictions) {Predictions};

@@

\node at (-1.5, 1.75) (metaprobetargs) {Probe targets};
\node at (-1.5, 1) (metaD2targs) {
%\(z^{prb}_{in}\)};
\(\left\{
\begin{matrix}
z^{func,prb}_{targ,0}\\
$\vdots$
\end{matrix}\right\}\)};

\node [align=center] at (0, 2.25) (metadispatch) {Loss (train)};

\path [arrow, dashed, out=180, in=-90] ([xshift=3]metaoutputs.west) to (metadispatch.south);

\path [arrow, dashed, out=0, in=-90] ([xshift=1]metaD2targs.east) to (metadispatch.south);

\end{scope}
\path [arrow, very thick, draw=bpurp, dotted, out=-40, in=-140] ([xshift=-1, yshift=3]zfunc.south) to ([xshift=19, yshift=3]metaD1.west);

%% evaluating meta mapping
\begin{scope}[shift={(11.8, 0)}]
\draw[boundingbox, draw=bblue, fill=white] (-2.25, -4.3) rectangle (2.5, 3.2);
\node[bblue] at (-0.35, 2.9) {Meta-mapping evaluation};

@@

\node at (-1, -1) (metaevalzfunc) {\(z^{func,pred}\)};

\node[block] at (0, -0.55) (metaevalH) {\(\mathcal{H}\)};
\path [arrow, out=45, in=180] ([yshift=-3]metaevalzfunc.north) to (metaevalH.west);

\node [block, dashed] at (1.5, -0.55) (metaevalF) {\(F_{z^{func,pred}}\)};
\path[arrow] (metaevalH.east) to (metaevalF.west);

@@

\node at (1.5, 1) (metaevaloutputs) {
%\(z^{pred}_{out}\)};
\(\left\{
\begin{matrix}
z^{pred}_{out,0},\\
$\vdots$
\end{matrix}\right\}\)};
\node at (1.5, 1.75) (metaevalpredictions) {Predictions};

@@

\node at (-1, 1.75) (metaevalprobetargs) {Probe targets};
\node at (-1, 1) (metaevalD2targs) {
%\(z^{prb}_{in}\)};
\(\left\{
\begin{matrix}
z^{prb}_{targ,0}\\
$\vdots$
\end{matrix}\right\}\)};

\node [align=center] at (0.3, 2.25) (metaevaldispatch) {Loss (eval)};

\path [arrow, dashed, out=180, in=-90] ([xshift=3]metaevaloutputs.west) to (metaevaldispatch.south);

\path [arrow, dashed, out=0, in=-90] ([xshift=-0.5]metaevalD2targs.east) to (metaevaldispatch.south);

@@

\caption{The HoMM architecture allows for transformations at different levels of abstraction. Basic meta-learning proceeds as in the gray box: a dataset consisting of (input, output) tuples is processed by the meta-network \(\mathcal{M}\) to produce a function embedding \(z^{func}\), which is processed by the hyper network \(\mathcal{H}\) to parameterize a function \(F_{z^{func}}\), which attempts to compute the transformation on held-out probe inputs. However, our approach goes beyond this basic meta-learning step. The function embedding \(z^{func}\) can then be seen as a single input or output at the next level of abstraction, when the same networks \(\mathcal{M}\) and \(\mathcal{H}\) are used to transform function embeddings based on examples of a meta-mapping (purple box). To evaluate meta-mapping performance, a probe embedding from a held-out funciton is transformed by the architecture to yield a predicted embedding for the transformed task. The performance of this predicted embedding is evaluated by moving back down a level of abstraction and evaluating on the actual target task (blue box). Because the function embedding is predicted by a transformation rather than computed from examples, this allows for zero-shot performance of a new task. (Blocks with solid edges denote learnable deep networks, and $F_{z}$ is a deep network parameterized by $\mathcal{H}$ conditioned on function embedding \(z\). See the text and appendix \ref{app_model_details} for details.)} \label{architecture_inference_fig}

@@

More formally, we treat functions and data as entities of the same type. From this perspective, the data points that one function receives can themselves be functions\footnote{Indeed, any data point can be represented as a constant function that outputs the data point.}. The key insight is that then our architecture can transform data points\footnote{Where ``data'' is a quite flexible term. The approach is agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.} to perform basic tasks (as is standard in machine learning), but it can also transform these task functions to adapt to new tasks. This is related to the idea of homoiconicity, defined above, and the concept of higher-order functions. Under this perspective, basic tasks and meta-mappings from task to task are really the same type of problem. The functions at one level of abstraction (the basic tasks) become inputs and outputs for higher-level functions at the next level of abstraction (meta-mapping between tasks). \par
Specifically, we embed each input, target, or mapping into a shared representational space $Z$. This means that single data points are embedded in the same space as the representation of a function or an entire dataset. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Model outputs are decoded from $Z$ by $\mathcal{O}: Z \rightarrow \text{output}$. Target outputs are encoded by $\mathcal{T}: \text{targets} \rightarrow Z$.\par

@@

More explicitly, suppose we have a dataset of example input, target pairs ($D_1 = \{(x_0, y_0), ...\}$), and some input $x$ from a probe dataset $D_2$. The system would predict a corresponding output $\hat{y}$ as: 

@@

where $F_{z^{func}}$ is the meta-learner's representation for the function underlying the examples in $D_1$:

@@

Alternatively, we could map from language to a meta-mapping embedding, rather than inducing it from examples of the meta-mapping. This corresponds to the human ability to change behavior in response to instructions. The key feature of our architecture -- the fact that tasks, data, and language are all embedded in a shared space -- allows for substantial flexibility within a unified framework. Furthermore, our approach is parsimonious. Because it uses the same meta-learner for both basic tasks and meta-mappings, this increased flexibility does not require any added parameters.\footnote{At least in principle, in practice of course increasing network size might be more beneficial for HoMM architectures performing meta-mappings as well as basic tasks than for those performing only basic tasks.}

%\subsection{Boolean functions}
%As a proof of concept, we first evaluated the system on a simple task of computing boolean functions on boolean inputs. Specifically, we can imagine mapping from binary-valued input vectors to a single binary output representing the evaluation of some logical proposition on the input. We train the system to figure out how to compute the predicate from seeing a subset of the (input, output) pairs. \par   
%We can then train the system to do various meta-mappings, for example identifying certain types of functions (like XOR) or negating predicates. \par
%
\section{Learning multivariate polynomials} \label{sec_poly}
\vspace{-0.5em} % section
\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/basic_results.png}
\caption{The polynomials domain, section \ref{sec_poly}.}
\label{poly_basic_results}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning.png}
\caption{The cards domain, section \ref{sec_cards}.}
\label{cards_basic_results}
\end{subfigure}%
\caption{The HoMM system succeeds at the basic meta-learning tasks, which are a necessary prerequisite for meta-mappings. (\subref{poly_basic_results}) Basic meta-learning in the polynomials domain, section \ref{sec_poly}. The system successfully generalizes to held out polynomials. The solid line indicates optimal performance; the dashed line indicates chance performance (an untrained model). (\subref{cards_basic_results}) Basic meta-learning in the cards domain, section \ref{sec_cards}. The system successfully generalizes to held out games, both when trained on a random sample of half the tasks, or when a targeted subset is held out. The gray dashed line indicates chance performance, while the solid lines are optimal performance. The orange dashed lines show how well the system could perform by playing the strategy from the most correlated trained task. The fact that it generally exceeds this difficult baseline shows a deeper form of generalization than just memorizing a few strategies and picking the closest. Error-bars throughout are bootstrap 95\%-confidence intervals, numerical values for plots can be found in appendix \ref{app_numerical_results}.}
\end{figure}
